Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-02-05 article/article.Rmd
+ replace adegenet with mclust (reverted from commit 08f8bcca6079915141bf3f8b3b3510f1d1a4edb9)

## Methods
### Study system
To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details]. This dataset was particularly well suited; it provides spatially explicit genomic data for a multitude of species across a broad geographic area over large environmental gradients. As part of this project, data were collected for 27 alpine plant species were across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1). Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell. Cells that did not contain any plant samples were omitted from our analysis. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453]. Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species. Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell. We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation. The majority of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

### Landscape data
We aimed to investigate the effectiveness of climatic and geographic surrogates. To describe the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted its two-dimensional coordinates. To describe the climatic variation among each cell, we obtained 19 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; using ArcMap version 10.2.2; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2). To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @r472).

@@

First, we grouped individuals into genetic lineages for each species. The purpose of this step was to avoid misclassifying loci as adaptive. We used Gaussian mixture models to assign individuals to populations [implemented in $\texttt{mclust}$ R package; @r477; @r467]. Briefly, for each species, we fit a suite of models to estimate the probability that individuals belonged different populations (using default priors and covariance structures). For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G} = 1--`max(r mclust.params.LST[[MODE]]$G)`). We identified the most parsimonious models (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used these to group individuals into populations (Figure S5).

We identified loci under selection. For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; Table S2). To avoid false-positive, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the $\texttt{Bayescan}$ manual). Additionally, to prevent individuals with unclear ancestry from biasing this analysis, we omitted individuals if there was considerable uncertainty in their population membership (maximum probability $\leq `r mclust.params.LST[[MODE]]$probability.threshold`$ % ). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32). For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457]. These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for all models; Table S3). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to summarize the typical genetic characteristics of individuals in the cells.

>>>>

## Methods
### Study system
To address the aims of the study, we utilized species distribution and genomic data collected by the IntraBioDiv project [@r451; see @r452 for methodological details]. This dataset was particularly well suited; it provides spatially explicit genomic data for a multitude of species across a broad geographic area over large environmental gradients. As part of this project, data were collected for 27 alpine plant species were across the European Alps according to a grid (22.3 km $\times$ 25 km; Figure S1). Briefly, in every second cell, plant samples were collected from three individuals for each species present inside the cell. Cells that did not contain any plant samples were omitted from our analysis. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453]. Matrices denoting the presence/absence of polymorphisms at loci were constructed independently for each species. Thus for each second grid cell, for each of the species in the grid cell, the dataset contains information describing the genomic variation of individuals inside the grid cell. We used these grid cells as planning units to develop prioritizations to investigate the effectiveness of surrogates for the (AFLP) genetic variation. Most of our analyses described hereafter were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

### Landscape data
We aimed to investigate the effectiveness of climatic and geographic surrogates. To describe the location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted its two-dimensional coordinates. To describe the climatic variation among each cell, we obtained 19 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principle components analysis (PCA; Table S1). The first two principle components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2). To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; GRUMP V1; @r472).

@@

First, we grouped individuals into genetic lineages for each species. The purpose of this step was to avoid misclassifying loci as adaptive. We used discriminant analyses of principle components [DAPC; @r467] to assign individuals to populations [implemented in $\texttt{adegenet}$ R package; @r468]. Briefly, for each species, we transformed the binary loci data using a principle components analysis (PCA) and obtained the principle components required to explain at least `r adegenet.params.LST[[MODE]]$perc.pca` % variation. We subject these components to multiple \textit{k}-means clustering models with varying numbers of assumed populations (\textit{k} = 1--`r adegenet.params.LST[[MODE]]$max.n.clust`; `r adegenet.params.LST[[MODE]]$n.iter` iterations; `r adegenet.params.LST[[MODE]]$n.start` random starts). We then  applied Ward's clustering method to the Bayesian Information Criterion (BIC) statistics associated with each model to partition the models into two groups, and selected the model before the first group switch (as recommended in the \texttt{adegenet} manual; Figure S4). The populations memberships associated with these models were used for subsequent analysis (Figure S5).

We identified loci under selection for each species (Table S2). For each species, we ran $\texttt{Bayescan}$ [version 2.1; @r456] using the population memberships identified in the previous analysis (1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations). We omitted loci where the global frequency of the minor allele was $\geq 0.05$ to avoid false-positives (following guidelines in the $\texttt{Bayescan}$ manual). To ensure convergence, we ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR; $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32). For each species, for each type of genetic variation detected, we used Gower distances [@r464] to express the genomic differences between individuals [using the $\texttt{cluster}$ R package to accommodate sparsity; @r457]. These distances were subject to multiple non-metric multi-dimensional scaling analyses [NMDS; implemented in the $\texttt{vegan}$ R package; @r458] using the minimum number of dimensions required to achieve an adequate stress value  ($\leq$ `r mds.params.LST[[MODE]]$max.stress`; `r mds.params.LST[[MODE]]$trymax` random starts for all models; Table S3). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the samples in each grid cell to characterize their the genetic variation of individuals in the cells.

### 2016-02-05 article/supporting_information.Rmd
+ replace adegenet with mclust (reverted from commit 08f8bcca6079915141bf3f8b3b3510f1d1a4edb9)

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=10, fig.cap='Relative support ($BIC$) for the number of populations in each species. Each panel denotes a different species. Each line denotes a different model parametrization. Asterisks indicate the model and number of populations with the most support.'}
# format data
bic.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		curr.MTX <- spp.Mclust.LST[[i]]$BIC
		attr(curr.MTX, 'dim') <- c(nrow(curr.MTX), ncol(curr.MTX))
		attr(curr.MTX, 'dimnames') <- list(
			seq_len(nrow(curr.MTX)),
			attr(curr.MTX, 'modelNames')
		)
		class(curr.MTX) <- 'matrix'
		curr.DF <- curr.MTX %>% as.data.frame %>% 
			mutate(G=attr(curr.MTX, 'G')) %>% 
			gather(Model,BIC,-G)
		curr.DF$BIC=curr.DF$BIC*-1
		return(
			mutate(
				curr.DF,
				Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE),
				best_G=curr.DF[which.min(curr.DF$BIC),'G'][[1]],
				best_stat=min(curr.DF$BIC,na.rm=TRUE),
				asterisk_col=c('black', rep(NA_character_, length(Species)-1))
			)
		)
	}
)

# make plot
suppressWarnings({
	ggplot(data=bic.DF) +
		geom_point(aes(y=BIC, x=G, color=Model)) +
		geom_line(aes(y=BIC, x=G, color=Model)) +
		geom_point(aes(y=best_stat, x=best_G), size=4, shape=8) +
		theme_classic() +
		scale_y_reverse()+
		theme(
			strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white', face='italic', size=20)
		) +
		xlab('Number of populations') +
		ylab('BIC') +
		facet_wrap(~ Species, ncol=4, scales='free_y')
})
```

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9.5, fig.height=10, fig.cap='Relative support ($BIC$) for the number of populations in each species. Each panel denotes a different species. Asterisks indicate the number of populations with the most support (see text for methodological details).'}
# format data
bic.buff <- laply(spp.Adegenet.LST, `[[`, 'Kstat') %>%
	c() %>% range() %>% diff %>% `*`(0.05)
bic.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		data.frame(
			k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$Kstat), fixed=TRUE)),
			stat=unname(spp.Adegenet.LST[[i]]$Kstat),
			best_k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$stat), fixed=TRUE)),
			best_stat=unname(spp.Adegenet.LST[[i]]$stat) + bic.buff,
			asterisk_col=c('black',rep('transparent', length(spp.Adegenet.LST[[i]]$Kstat)-1)),
			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
		)
	}
)
# make plot
suppressWarnings({
	ggplot(data=bic.DF) +
		geom_point(aes(y=stat, x=k)) +
		geom_point(aes(y=best_stat, x=best_k), size=4, shape=8) +
		geom_line(aes(y=stat, x=k)) +
		theme_classic() +
		theme(
			strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white', face='italic', size=20)
		) +
		xlab('Number of populations') +
		ylab('BIC') +
		facet_wrap(~ Species, ncol=4)
})
```
